---
  output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

## Overview 
pmparams helps define and format parameter tables.

An example parameter table script highlighting `pmparams` functions to create a parameter table can be found
[here](https://github.com/metrumresearchgroup/expo1-nonmem-foce/blob/main/script/pk-final-model-table.R).

An example parameter table script highlighting `pmparams` functions to create a bootstrap parameter table can be found
[here](https://github.com/metrumresearchgroup/expo1-nonmem-foce/blob/main/script/pk-final-model-table-boot.R).

## Documentation
Public documentation of all functions is hosted at [https://metrumresearchgroup.github.io/pmparams/](https://metrumresearchgroup.github.io/pmparams/)


## Development

`pmparams` uses [pkgr](https://github.com/metrumresearchgroup/pkgr) to manage
development dependencies and [renv](https://rstudio.github.io/renv/) to
provide isolation. To replicate this environment,

1.  clone the repo

2.  install pkgr

3.  open package in an R session and run `renv::init(bare = TRUE)`

    -   install `renv` \> 0.8.3-4 into default `.libPaths()` if not
        already installed

4.  run `pkgr install` in terminal within package directory

5.  restart session

Then, launch R with the repo as the working directory (open the project
in RStudio). renv will activate and find the project library.

## Getting help

If you encounter a clear bug, please file an issue with a minimal reproducible example on [pmparams](https://github.com/pmparams/issues). 

```{r setup, include = FALSE}
## Package coverage
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
if (basename(getwd()) == "docs") {
  knitr::opts_knit$set(root.dir = file.path(getwd(), ".."))
}
```

```{r, include=FALSE}
# covr::package_coverage(type = "all", 
#                        line_exclusions = c("R/print-aesthetics.R", "R/list-files-of-type.R", 
#                                            "R/find-in-files.R", "R/gather-flags.R"))
```
