---
title: "Getting Started with pmparams"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{getting-started}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

options(
  'bbr.bbi_exe_path' = bbr::read_bbi_path(),
  'bbr.verbose' = FALSE
)

# NOTE: if running chunks interactively we need to load the package first
devtools::load_all()

# bbr required to render some sections of this vignette
# - used for CI
eval_bbr_chunks <- is.null(check_for_suggested_pkgs("bbr"))

# CI has issues running st_as_image and stable_save_image
# - Only build this vignette via pkgdown or when rendering manually (must be on metworx)
render_tables <- (Sys.getenv("METWORX_VERSION") != "")
render_bbr_tables <- render_tables && eval_bbr_chunks
```


# Introduction

This vignette takes the user through some basic scenarios for defining, formatting, and making parameter tables using `pmparams`. For more information how to effectively integrate `pmparams` into your workflow, visit the 
<a href="https://merge.metrumrg.com/expo/expo1-nonmem-foce/posts/parameter-tables.html" target="_blank">
MeRGE Expo: Parameter Tables </a>.

```{r load packages, results = 'hide', message=FALSE, warning=FALSE}
library(pmparams)
library(bbr)
library(dplyr)
library(pmtables)
```


# Data requirements

## Parameter key

We begin by creating a parameter key that tells R how to interpret you parameter values. Our code require four arguments for each parameter:

* abb - abbreviation for model parameter (we use latex coding)
* desc - parameter description to appear
* panel - the table panel the parameter should appear under
* trans - definition of how the parameter should be transformed

If you have a model that uses theta in the $ERROR block, make sure that the theta section's panel is "RV" so that `pmparams` functions can properly identify it. 

It is recommended to use a parameter key yaml, but `pmparams` works for parameter key tibbles. 

A more detailed walk-through of generating the parameter key is available here: <a href="https://merge.metrumrg.com/expo/expo1-nonmem-foce/posts/parameter-key.html" target="_blank">
MeRGE Expo: Creating a Parameter Key </a>. 

```{r, message = FALSE}
model_dir <- system.file("model/nonmem", package = "pmparams")
paramKey <- file.path(model_dir, "pk-parameter-key-new.yaml")

yaml::yaml.load_file(paramKey) %>% unlist() %>% head()
```

## Parameter estimates

`pmparams` allows for different parameter estimate input types:

* path to a `bbr` NONMEM model 
* `bbr` NONMEM model (output from `bbr::read_model()`)
* `bbr` NONMEM model summary (output from `bbr::model_summary()`)
* data.frame of parameter estimates (e.g., output from `bbr::param_estimates()`)

For this example, we will use a `bbr` model
```{r, message = FALSE, eval = eval_bbr_chunks}
mod <- bbr::read_model(file.path(model_dir, "102"))
```


# Define parameter table

We will now join parameter estimates and parameter key. Note: this is an `inner_join`, so only parameters included in the model output and parameter key will be kept in the table. This was done so that, if your base and final model used the same structural THETAs and random parameters, the same parameter key could be used for both. The additional covariate THETAs defined in the parameter key YAML would simply be ignored when creating the base model parameter table.

Additionally, `define_param_table` performs checks and calculates confidence intervals.

```{r, eval = eval_bbr_chunks}
param_df <- mod %>% 
  define_param_table(.key = paramKey)

head(param_df)
```

# Format parameter table

Now, we perform some house-keeping based on the new parameter key information, calculate any summary statistics (the 95% confidence intervals are calculated by default), and format the values for the report using `format_param_table()`.
```{r, eval = eval_bbr_chunks}
param_df <- param_df %>% 
  format_param_table(.cleanup_cols = TRUE, .digit = 3)

head(param_df)
```

Usually `define_param_table` and `format_param_table` are run together in a single call like this:

```{r, eval = FALSE}
param_df <- define_param_table(
  .estimates = paramEstimatePath,
  .key = paramKeyPath
) %>% format_param_table()
```

# Make parameter table

We can leverage `make_pmtable()` to generate fixed effects and random effects parameter tables, and `param_notes()` to easily add informative footnotes or notes.

```{r}
# grab footnotes
footnote <- param_notes(.ci = 95)
```

```{r, eval = render_bbr_tables}
##  FIXED EFFECTS table ----------------------------
tab_fixed <- make_pmtable(param_df, .pmtype = "fixed") %>% 
  # abbreviations
  st_notes(footnote$ci, footnote$se) %>% 
  st_notes_str() %>% 
  # equations
  st_notes(footnote$ciEq) %>%
  # file
  st_files(output = "deliv/final-param-fixed-ci-95.tex")

tab_fixed_out <- tab_fixed %>% stable() %>% 
  stable_save_image(format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_fixed_out, width = 0.6)
```


```{r, eval = render_bbr_tables}
##  RANDOM EFFECTS table ----------------------------
tab_random <- make_pmtable(param_df, .pmtype = "random") %>% 
  # abbreviations
  st_notes(footnote$ci, footnote$se) %>% 
  st_notes_str() %>% 
  # equations
  st_notes(footnote$cvOmegaEq, footnote$cvSigmaEq) %>%
  # file
  st_files(output = "deliv/final-param-random-ci-95.tex")

tab_random_out <- tab_random %>% stable() %>% 
  stable_save_image(format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_random_out, width = 0.6)
```

If you want to add or overwrite pre-defined `pmtables` argument, simply pipe onto the `make_pmtable` tibble:
```{r, eval = render_bbr_tables}
tab_random2 <- tab_random %>% 
  st_panel("abb") %>% 
  st_select(-"type", -"greek") %>% 
  st_files(output = "deliv/final-param-random-ci-95-modified.tex")

tab_random_out2 <- tab_random2 %>% stable() %>% 
  stable_save_image(format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_random_out2, width = 0.6)
```

# Other functions
`pmparams` exports two functions for appending bootstrap estimates to parameter tables. See the 
<a href="https://metrumresearchgroup.github.io/pmparams/vignettes/making-bootstrap-parameter-tables.html" 
target="_blank">Reference Log</a> for more details.
