---
title: "Making Bootstrap Parameter Tables"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{making-bootstrap-parameter-tables}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---
  
  
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

options(
  'bbr.bbi_exe_path' = bbr::read_bbi_path(),
  'bbr.verbose' = FALSE
)

# NOTE: if running chunks interactively we need to load the package first
devtools::load_all()

# bbr > 1.11.0 is required to render most sections of this vignette
# - used for CI
bbr_installed <- is.null(check_for_suggested_pkgs("bbr"))
eval_bbr_chunks <- bbr_installed && version_above("bbr", "1.11.0")

# CI has issues running st_as_image and stable_save_image
# - Only build this vignette via pkgdown or when rendering manually (must be on metworx)
render_tables <- (Sys.getenv("METWORX_VERSION") != "")
render_bbr_tables <- render_tables && eval_bbr_chunks
```

# Introduction

This is a vignette to help you become familiar with using `define_boot_table()` and `format_boot_table()`. 
For more information how to effectively integrate `pmparams` into your workflow, visit the 
<a href="https://merge.metrumrg.com/expo/expo1-nonmem-foce/posts/parameter-tables.html" target="_blank">
MeRGE Expo: Parameter Tables</a>.

```{r load packages, results = 'hide', message=FALSE, warning=FALSE}
library(pmparams)
library(bbr)
library(dplyr)
library(yaml)
library(pmtables)
```

# Data requirements

Before we get started, lets set some key variables we'll use throughout this tutorial, including the original `bbr` model and the bootstrap run:
```{r}
# Model Directory
model_dir <- system.file("model/nonmem", package = "pmparams")
# Parameter Key
param_key <- file.path(model_dir, "pk-parameter-key-new.yaml")
# Model we're bootstrapping
orig_mod <- bbr::read_model(file.path(model_dir, "106"))
# Bootstrap model object
boot_run <- bbr::read_model(file.path(model_dir, "106-boot"))
```

## Parameter key

We begin by creating a parameter key that tells R how to interpret your parameter values. Our code requires four arguments for each parameter:

* abb - abbreviation for model parameter (we use latex coding)
* desc - parameter description to appear
* panel - the table panel the parameter should appear under
* trans - definition of how the parameter should be transformed

If you have a model that uses theta in the $ERROR block, make sure the theta panel label is "RV" so the `pmparams` functions can properly identify it. 

We recommend using a parameter key yaml, but `pmparams` works for parameter key tibbles. 

A more detailed walk-through of generating the parameter key is available here: <a href="https://merge.metrumrg.com/expo/expo1-nonmem-foce/posts/parameter-key.html" target="_blank">
MeRGE Expo: Creating a Parameter Key </a>. 

```{r, message = FALSE}
param_key_lst <- yaml::yaml.load_file(param_key)
head(unlist(param_key_lst))
```

# Creating a parameter table bootstrap estimates

## Define & format bootstrap parameter table
Parameter tables of bootstrap estimates are created by first passing your bootstrap estimates to the `pmparams` `define_boot_table()` function, which handles both the joining and creation of new confidence interval related columns.

- Note: this uses `dplyr::inner_join`, so only parameters included in the model output _and_ parameter key will be kept in the table. This was done so that if your base and final model used the same structural THETAs and random parameters, the same parameter key could be used for both. The additional covariate THETAs defined in the parameter key YAML would simply be ignored when creating the base model parameter table.

Bootstrap runs can be performed directly in `bbr` (version `1.11.0` or later), which tracks a bootstrap model object (`boot_run`). Once executed, the `bbr::bootstrap_estimates()` function can be used to easily read those estimates into a data frame that you can subsequently pass to the `pmparams` `define_boot_table()` function. 

### Bootstrap parameter estimates argument
To support use cases outside of `bbr`, this function allows you to pass in those bootstrap estimates in a couple of formats:

* A ***path*** to a csv (`file.path(model_dir, "boot/data/boot-106.csv")`)
* A data.frame (like what's returned from `bbr::bootstrap_estimates()`)

```{r, eval = eval_bbr_chunks}
# Also accepted
 # boot_ests <- file.path(model_dir, "boot/data/boot-106.csv")
 # boot_ests <- readr::read_csv(file.path(model_dir, "boot/data/boot-106.csv"))
boot_ests <- bbr::bootstrap_estimates(boot_run)

boot_df <- define_boot_table(
  .boot_estimates = boot_ests,
  .key = param_key
)

head(boot_df)
```

Now we perform some house-keeping based on the new parameter key information and format the values for the report using `format_boot_table()`.
```{r, eval = eval_bbr_chunks}
boot_df <- format_boot_table(boot_df)

head(boot_df)
```


# Combine bootstrap and original model parameter estimates

Start by creating a parameter table for your _original model_, and then join it to the formatted bootstrap table (`boot_df`) we created above:
```{r, eval = eval_bbr_chunks}
param_df <- orig_mod %>%
  define_param_table(param_key, .ci = 95) %>% 
  format_param_table()


boot_param_df <- left_join(param_df, boot_df, by = c("abb", "desc"))
head(boot_param_df)
```


```{r, eval = render_bbr_tables}
# Define bootstrap CI footnote
footBoot <- boot_notes(.ci = 95, .n_run = nrow(boot_ests))

# Render table
tab <- boot_param_df %>%
  make_boot_pmtable(.pmtype = "fixed") %>%
  st_center(desc = col_ragged(5.5), abb = "l") %>% 
  st_notes(param_notes()$ci) %>%                   # add abbreviations
  st_notes_str() %>%                               # collapse all abbreviations to a string                        
  st_notes(param_notes()$logTrans, footBoot$boot_ci) %>%   # customize other notes
  st_notes_detach(width = 1) %>% 
  st_files(output = "final-param-fixed-boot.tex") %>%
  stable(output_file = "deliv/mod106-boot-95-ci-fixed.tex")

tab_out <- stable_save_image(tab, format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_out, width = 0.8)
```


## Complex example
To illustrate the flexibility of `make_boot_pmtable()`, we'll use a more complex example where we include two confidence intervals:
```{r, eval = eval_bbr_chunks, message=FALSE}
# Two confidence intervals
boot_df <- boot_ests %>%
  define_boot_table(param_key, .ci = 95) %>% 
  format_boot_table()

boot_df2 <- boot_ests %>%
  define_boot_table(param_key, .ci = 90) %>% 
  format_boot_table()


boot_param_df <- param_df %>% 
  left_join(boot_df) %>% 
  left_join(boot_df2)
```

```{r, eval = render_bbr_tables}
tab <- boot_param_df %>%
  make_boot_pmtable(.pmtype = "fixed") %>% 
  st_center(desc = col_ragged(5.5), abb = "l") %>% 
  st_notes(param_notes()$ci) %>%         # add abbreviations
  st_notes_str() %>%                     # collapse all abbreviations to a string                        
  st_notes(param_notes()$logTrans) %>%   # customize other notes
  st_notes_detach(width = 1) %>% 
  st_files(output = "final-param-fixed-boot-multi-ci.tex") %>% 
  stable(output_file = "deliv/mod106-boot-95-90-ci-fixed.tex")

tab_out <- stable_save_image(tab, format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_out, width = 0.8)
```


