---
title: "Making Bootstrap Parameter Tables"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{making-bootstrap-parameter-tables}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---
  
  
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")

options(
  'bbr.bbi_exe_path' = bbr::read_bbi_path(),
  'bbr.verbose' = FALSE
)

# NOTE: if running chunks interactively we need to load the package first
devtools::load_all()

# bbr > 1.11.0 is required to render most sections of this vignette
# - used for CI
bbr_installed <- is.null(check_for_suggested_pkgs("bbr"))
eval_bbr_chunks <- bbr_installed && version_above("bbr", "1.11.0")

# CI has issues running st_as_image and stable_save_image
# - Only build this vignette via pkgdown or when rendering manually (must be on metworx)
render_tables <- (Sys.getenv("METWORX_VERSION") != "")
render_bbr_tables <- render_tables && eval_bbr_chunks
```

# Introduction

This is a vignette to help you become familiar with using `define_boot_table()` and `format_boot_table()`. 
For more information how to effectively integrate `pmparams` into your workflow, visit the 
<a href="https://merge.metrumrg.com/expo/expo1-nonmem-foce/posts/parameter-tables.html" target="_blank">
MeRGE Expo: Parameter Tables </a>..

```{r load packages, results = 'hide', message=FALSE, warning=FALSE}
library(pmparams)
library(bbr)
library(dplyr)
library(yaml)
library(pmtables)
```

# Data requirements

## Bootstrap methods

Bootstrapping a `bbr` model previously was a more involved process that required manually saving out csv. However as of `bbr 1.11.0`, bootstrapping can be done in-house, tracking a new bootstrap model object. For this reason we support multiple methods for specifying the estimates, all of which will be described below.

In any case, lets set some key variables we'll use throughout this tutorial:
```{r}
model_dir <- system.file("model/nonmem", package = "pmparams")
# Parameter Key
paramKeyPath <- file.path(model_dir, "pk-parameter-key-new.yaml")
# Model we're bootstrapping
orig_mod <- bbr::read_model(file.path(model_dir, "106"))
```

## Parameter key

We begin by creating a parameter key that tells R how to interpret you parameter values. Our code require four arguments for each parameter:

* abb - abbreviation for model parameter (we use latex coding)
* desc - parameter description to appear
* panel - the table panel the parameter should appear under
* trans - definition of how the parameter should be transformed

If you have a model that uses theta in the $ERROR block, make sure that the theta section's panel is "RV" so that `pmparams` functions can properly identify it. 

It is recommended to use a parameter key yaml, but `pmparams` works for parameter key tibbles. 

A more detailed walk-through of generating the parameter key is available here: <a href="https://merge.metrumrg.com/expo/expo1-nonmem-foce/posts/parameter-key.html" target="_blank">
MeRGE Expo: Creating a Parameter Key </a>. 

```{r, message = FALSE}
paramKey <- yaml::yaml.load_file(paramKeyPath)
head(unlist(paramKey))
```

# Creating Bootstrap estimate parameter table with `bbr 1.11.0` and greater

We begin by reading in the original `bbr` model and the bootstrap run:
```{r, eval = eval_bbr_chunks}
orig_mod <- bbr::read_model(file.path(model_dir, "106"))
boot_run <- bbr::read_model(file.path(model_dir, "106-boot"))
```

## Define & Format bootstrap parameter table
We will now join parameter estimates and parameter key. `define_param_table` will handle both the joining and creation of new confidence interval related columns.

- Note: this uses `dplyr::inner_join`, so only parameters included in the model output _and_ parameter key will be kept in the table. This was done so that if your base and final model used the same structural THETAs and random parameters, the same parameter key could be used for both. The additional covariate THETAs defined in the parameter key YAML would simply be ignored when creating the base model parameter table.

```{r, eval = eval_bbr_chunks}
boot_df <- define_boot_table(
  .boot_estimates = bbr::bootstrap_estimates(boot_run),
  .key = paramKeyPath
)

head(boot_df)
```

Now we perform some house-keeping based on the new parameter key information, and format the values for the report using `format_boot_table`.
```{r, eval = eval_bbr_chunks}
boot_df <- format_boot_table(boot_df)

head(boot_df)
```

# Creating Bootstrap estimate paramter table using previous methods

### Bootstrap parameter estimates 
`pmparams` allows for different boot parameter estimate input types:
* path to a csv 
* A data.frame (like what's returned from `bbr::bootstrap_estimates()`)

## Define & Format bootstrap parameter table

For this example, we will use the path to a csv:
```{r, message = FALSE}
boot_ests_path <- file.path(model_dir, "boot/data/boot-106.csv")
```

As we can see below, we get the same final data frame as before:
```{r}
# Join to key and calculate confidence interval
boot_df <- define_boot_table(
  .boot_estimates = boot_ests_path,
  .key = paramKeyPath
)

# Format table
boot_df <- format_boot_table(boot_df)

head(boot_df)
```


# Make a simple bootstrap parameter table

Below is a simple display of the bootstrap confidence intervals using well known `pmtables` functions. As we'll see a little later, we can use `make_boot_pmtable()` to make a much nicer table.
```{r, eval = render_tables}
# Grab footnotes
footnote <- param_notes(.ci = 95)

# Render table
tab <- boot_df %>%
  st_new() %>% 
  st_panel("desc") %>% 
  st_span("Non-parametric bootstrap", c("boot_median", "boot_ci_95")) %>%
  st_rename(
    Median = boot_median,
    "95\\% CI" = boot_ci_95,
    " "= abb
  ) %>% 
  # abbreviations
  st_notes(footnote$ci) %>% 
  # equations
  st_notes(footnote$ciEq) %>%
  stable(output_file = "deliv/simple-boot.tex")

tab_out <- stable_save_image(tab, format = "png")
```

```{r, eval = render_tables, fig.align='center'}
st_image_show(tab_out, width = 0.5)
```


# Combine bootstrap and original model parameter estimates

Start by creating both of the formatted parameter tables, and then join them together as seen below:
```{r, eval = eval_bbr_chunks}
boot_df <- define_boot_table(
  .boot_estimates = bbr::bootstrap_estimates(boot_run),
  .key = paramKeyPath,
  .ci = "iqr"
) %>% format_boot_table()

param_df <- define_param_table(
  .estimates = orig_mod,
  .key = paramKeyPath,
  .ci = 95
) %>% format_param_table()


boot_param_df <- left_join(param_df, boot_df, by = c("abb", "desc"))
head(boot_param_df)
```


```{r, eval = render_bbr_tables}
# Define footnote
footBoot <- glue::glue("CI were determined from the 2.5th and 97.5th percentiles 
                of the non-parametric bootstrap estimates (n ={nrow(read.csv(boot_ests_path))}).")

# Render table
tab <- make_boot_pmtable(.df = boot_param_df, .pmtype = "fixed") %>%
  st_center(desc = col_ragged(5.5), abb = "l") %>% 
  st_notes(param_notes()$ci) %>%                   # add abbreviations
  st_notes_str() %>%                               # collapse all abbreviations to a string                        
  st_notes(param_notes()$logTrans, footBoot) %>%   # customize other notes
  st_notes_detach(width = 1) %>% 
  st_files(output = "final-param-fixed-boot.tex") %>%
  stable(output_file = "deliv/mod106-boot-95-ci-fixed.tex")

tab_out <- stable_save_image(tab, format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_out, width = 0.5)
```


## Complex example
To illustrate the flexibility of `make_boot_pmtable()`, we'll use a more complex example. Here we'll specify `.percentiles` to include two confidence intervals, the median value, and an extra 25% quantile (just for fun).

```{r, eval = eval_bbr_chunks}
boot_df <- define_boot_table(
  .boot_estimates = bbr::bootstrap_estimates(boot_run),
  .key = paramKeyPath,
  .percentiles = c(0.025, 0.05, 0.25, 0.5, 0.95, 0.975)
) %>% format_boot_table()


boot_param_df <- left_join(param_df, boot_df, by = c("abb", "desc"))
```

```{r, eval = render_bbr_tables}
##  FIXED EFFECTS table ----------------------------
tab <- make_boot_pmtable(.df = boot_param_df, .pmtype = "fixed") %>% 
  st_center(desc = col_ragged(5.5), abb = "l") %>% 
  st_notes(param_notes()$ci) %>%                   # add abbreviations
  st_notes_str() %>%                               # collapse all abbreviations to a string                        
  st_notes(param_notes()$logTrans) %>%   # customize other notes
  st_notes_detach(width = 1) %>% 
  st_files(output = "final-param-fixed-boot-multi-ci.tex") %>% 
  stable(output_file = "deliv/mod106-boot-95-90-ci-25-perc-fixed.tex")

tab_out <- stable_save_image(tab, format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_out, width = 0.5)
```

```{r, eval = render_bbr_tables}
##  RANDOM EFFECTS table ----------------------------
tab <- make_boot_pmtable(.df = boot_param_df, .pmtype = "random") %>% 
  st_center(desc = col_ragged(5.5), abb = "l") %>% 
  st_notes(param_notes()$ci) %>%                   # add abbreviations
  st_notes_str() %>%                               # collapse all abbreviations to a string                        
  st_notes(param_notes()$logTrans) %>%   # customize other notes
  st_notes_detach(width = 1) %>% 
  st_files(output = "final-param-random-boot-multi-ci.tex") %>% 
  stable(output_file = "deliv/mod106-boot-95-90-ci-25-perc-random.tex")

tab_out <- stable_save_image(tab, format = "png")
```

```{r, eval = render_bbr_tables, fig.align='center'}
st_image_show(tab_out, width = 0.5)
```

